---
title: "Simulations"
author: "Thijs Carriere"
date: "1-12-2021"
output: html_document
---

## 0. Libraries & Goal 

```{r Needed packages, message=FALSE, warning=FALSE, comment=FALSE}
# Packages
library(tidyverse)
library(lavaan)
```

The goal of this simulation is to find data for the table 1 in the Research Report. The b values in the table are obtained in this simulation, where MIMIC models are run for different simulated test items.


## 1. Simulating Data
### 1.1 Obtaining Thetas

Making 4 groups with a difference in their latent trait. Group 3 has a mean of .75, Group 2 of .25, Group 1 of -.25 and group 4 of -1. Groups consist of 100 observations.

```{r}
# Making two groups with a difference in theta of .5

# Set seed
set.seed(100)

count.1 <- rnorm(100, -.25, 1)
count.2 <- rnorm(100, .25, 1)
count.3 <- rnorm(100, .75, 1)
count.4 <- rnorm(100, -1, 1)

thetas <- as.data.frame(c(count.1, count.2, count.3, count.4)) # combine them in a frame
colnames(thetas) <- "Theta"

# give them groups (countries)
thetas$group <- 1
thetas[101:200, 2] <- 2
thetas[201:300, 2] <- 3
thetas[301:400, 2] <- 4


# Generate IDs for observations
thetas$ID <- 1:400
```


### 1.2 Obtaining responses

Generating 6 scores with a 2PL model. Parameters obtained randomly, except for the 6th item. This item is more difficult for the group with the higher mean theta. Also, the item had a different discrimination parameter for non-uniform DIF.

```{r Obtaining Responses}
# parameters
set.seed(15)
b <- c(rnorm(5, 0, 1), .5, 1)
a <- c(runif(5, 1, 3), .5, 2)

# Last item will be harder for the second group, while they have higher theta on average. 

# Get responses
resp <- matrix(NA, nrow = 400, ncol = 6) # create storage
colnames(resp) <- c("item1", "item2", "item3", "item4", "item5", "item6")

# Responses by binomial 2 parameter model for item 1 to item 5
set.seed(113)
for(i in 1:400){
  for(j in 1:5){
    thet <- exp(a[j]*(thetas[i,1] - b[j]))/(1 + exp(a[j]*(thetas[i,1] - b[j])))
    resp[i,j] <- rbinom(1, 1, thet)
  }
}

# Do last item by separately to change the item parameters over groups
for(i in 1:200){
  thet <- exp(a[6]*(thetas[i,1] - b[6]))/(1 + exp(a[6]*(thetas[i,1] - b[6])))
    resp[i,6] <- rbinom(1, 1, thet)
}

for(i in 301:400){
  thet <- exp(a[6]*(thetas[i,1] - b[6]))/(1 + exp(a[6]*(thetas[i,1] - b[6])))
    resp[i,6] <- rbinom(1, 1, thet)
}

for(i in 201:300){
  thet <- exp(a[7]*(thetas[i,1] - b[7]))/(1 + exp(a[7]*(thetas[i,1] - b[7])))
    resp[i,6] <- rbinom(1, 1, thet)
}  


```


### 1.3 Generating response times

As process data response times are simulated. For the last item, the second group has average higher response times (40 and 80 seconds). 

```{r Response times}
# Add response times, make for 6 a bit different
times <- matrix(NA, 400, 6)
colnames(times) <- c("T1", "T2", "T3", "T4", "T5", "T6")

set.seed(100)
for(i in 1:5){
  mean <- runif(1, 20, 80)
  times[,i] <- rnorm(400, mean, sd = 15)
}

# for question 6, group 2 needs more time than group 1
times[1:200,6] <- rnorm(100, 40, 10)
times[301:400,6] <- rnorm(100, 40, 10)
times[201:300,6] <- rnorm(100, 80, 10)


# 1 big frame
data.set <- cbind(thetas, resp, times)
data.set <- data.set %>% select(ID, everything())
data.set$group <- as.factor(data.set$group)

head(data.set)
```

### 1.4 Making dummies

```{r}
# making dummies
data.set$dum1 <- ifelse(data.set$group == "1", "1", "0")
data.set$dum2 <- ifelse(data.set$group == "2", "1", "0")
data.set$dum3 <- ifelse(data.set$group == "3", "1", "0")
data.set$dum4 <- ifelse(data.set$group == "4", "1", "0")

data.set <- data.set %>% select(1:9, 16:19, 10:15)
```


## 2. Detecting uniform DIF

### 2.1 Reference model

```{r}
difdum <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod <- sem(model = difdum, data = data.set, ordered = colnames(data.set[,3:13]) )
summary(difdum.mod, fit.measures = T)
```

### 2.2 MIMIC models for 4 items

For the four items presented in the table of the research report, MIMIC models are run. Uniform DIF is investigated for each of these items. The obtained b-values are used in the table in the research report.

#### item 1

```{r warning=FALSE}
# item 1
difdum <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod <- sem(model = difdum, data = data.set, ordered = colnames(data.set[,3:13]) )

# dif with dummies
difdum_i1 <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
item1 ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod_i1 <- sem(model = difdum_i1, data = data.set, ordered = colnames(data.set[,3:13]) )
summary(difdum.mod_i1, fit.measures = T)

anova(difdum.mod, difdum.mod_i1)

```

#### item 2

```{r warning=FALSE}
# item 2
difdum <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod <- sem(model = difdum, data = data.set, ordered = colnames(data.set[,3:13]) )

# dif with dummies
difdum_i2 <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
item2 ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod_i2 <- sem(model = difdum_i2, data = data.set, ordered = colnames(data.set[,3:13]) )
summary(difdum.mod_i2, fit.measures = T)

anova(difdum.mod, difdum.mod_i2)
```

#### item 3

```{r warning=FALSE}
# item 3
difdum <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod <- sem(model = difdum, data = data.set, ordered = colnames(data.set[,3:13]) )

# dif with dummies
difdum_i3 <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
item3 ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod_i3 <- sem(model = difdum_i3, data = data.set, ordered = colnames(data.set[,3:13]) )
summary(difdum.mod_i3, fit.measures = T)

anova(difdum.mod, difdum.mod_i3)
```

#### item 6

```{r warning=FALSE}
# dif with dummies
difdum2 <- '
lat =~ item1 + item2 + item3 + item4 + item5 + item6 
lat ~ dum1 + dum2 + dum3 + dum4
item6 ~ dum1 + dum2 + dum3 + dum4
'

# Run model and summary
difdum.mod2 <- sem(model = difdum2, data = data.set, ordered = colnames(data.set[,3:13]) )
summary(difdum.mod2, fit.measures = T)

anova(difdum.mod, difdum.mod2)
```
