---
title: "Online TIMSS data"
author: "Thijs Carrière"
date: "November 2021"
output:
  html_document:
    df_print: paged
---
### 0. Goals of this document
In this document, the TIMSS data files for the used countries in this project are loaded and pre-processed. This means that the data sets are modified in such a way that they are useable for the analyses further in this project.

### 1. Working directionary and packages

First the needed packages are loaded

```{r warning=FALSE, message=FALSE}
# Needed packages
library(foreign)
library(tidyverse)
library(magrittr)
library(dexter)
library(plyr)
library(ggplot2)
library(mirt)
library(lavaan)
```


### 2. Reading in data and making list out of it

The data sets of the different countries are loaded in.

```{r warning = FALSE, message = FALSE}

# Loading in data of used countries 
timss_fra <- read.spss("../../2. data/2.1 Raw TIMSS data/bsafram7.sav", to.data.frame = T)
timss_eng <- read.spss("../../2. data/2.1 Raw TIMSS data/bsaengm7.sav", to.data.frame = T)
timss_norw <- read.spss("../../2. data/2.1 Raw TIMSS data/bsanorm7.sav", to.data.frame = T)
timss_hkg <- read.spss("../../2. data/2.1 Raw TIMSS data/bsahkgm7.sav", to.data.frame = T)
timss_qat <- read.spss("../../2. data/2.1 Raw TIMSS data/bsaqatm7.sav", to.data.frame = T)
timss_kor <- read.spss("../../2. data/2.1 Raw TIMSS data/bsakorm7.sav", to.data.frame = T)
timss_twn <- read.spss("../../2. data/2.1 Raw TIMSS data/bsatwnm7.sav", to.data.frame = T)

# Make list with dfs
timss <- list(timss_fra, timss_eng, timss_hkg, timss_kor, timss_qat, timss_norw, timss_twn)

```


### 3. Cleaning data and add scores

Here, items are scored and unneeded information is removed from the data sets. Information about the items used and the choices in scoring and removing can be found in the TIMSS userguide, page 63.

```{r message=FALSE, warning=FALSE}
### Data cleaning on List ###
timss <- list(timss_fra, timss_eng, timss_hkg, timss_kor, timss_qat, timss_norw, timss_twn)

# names to list
names(timss) <- c("France", "England", "Hongkong", "Korea", "Qatar", "Norway", "Taiwan")
names <- c("France", "England", "Hongkong", "Korea", "Qatar", "Norway", "Taiwan")

# subset list
timss <- lapply(timss, filter, IDBOOK == "Booklet 01")

# clear empty columns
for(i in 1:length(timss)){
  timss[[i]] <- Filter(function(x)!all(is.na(x)), timss[[i]])
}

# subset relevant columns
for(i in 1:length(timss)){
  timss[[i]] <- timss[[i]] %>% dplyr::select(-starts_with("SE"))
}

items <- colnames(timss$France[,6:49])

# creating dev branche
timss.speed.proto <- lapply(timss, dplyr::select, IDCNTRY:IDSTUD, ME52024_S:ME72209_S)
timss.freq.proto <- lapply(timss, dplyr::select, IDCNTRY:IDSTUD, ME52024_F:ME72209_F)

# pivot longer & relocate
timss <- lapply(timss, pivot_longer, all_of(items), names_to = "item", values_to = "response")
timss <- lapply(timss, relocate, c(item, response), .after = IDSTUD)

# Filter wrong items
timss <- lapply(timss, filter, item != "ME72007A" & item !="ME72002B" & item != "ME72007C" & item != "ME72007D" & item != "ME72007E" & item != "ME72180A" & item != "ME72180B" & item != "ME72180C" & item != "ME72198A" & item != "ME72198B" & item != "ME72170A" & item != "ME72170B" & item != "ME72170C")

# This one item won't go unless doing this way ...
timss$France <- filter(timss$France, item != "ME72007B")
timss$England <- filter(timss$England, item != "ME72007B")
timss$Hongkong <- filter(timss$Hongkong, item != "ME72007B")
timss$Korea <- filter(timss$Korea, item != "ME72007B")
timss$Qatar <- filter(timss$Qatar, item != "ME72007B")
timss$Norway <- filter(timss$Norway, item != "ME72007B")
timss$Taiwan <- filter(timss$Taiwan, item != "ME72007B")

# Select relevant variables
timss <- lapply(timss, dplyr::select, IDCNTRY:response, ITSEX)


# Make scoring
timss <- lapply(timss, mutate, item_score = case_when(
    response == "CORRECT RESPONSE" & item == "ME72007" ~ 2,
    response == "PARTIALLY CORRECT RESPONSE" & item == "ME72007" ~ 1,
    response == "INCORRECT RESPONSE" ~ 0,
    response == "CORRECT RESPONSE" ~ 1,
    response == "A*" ~ 1,
    response == "B*" ~ 1,
    response == "C*" ~ 1,
    response == "D*" ~ 1,
    response == "A" ~ 0,
    response == "B" ~ 0,
    response == "C" ~ 0,
    response == "D" ~ 0,
    response == "INCORRECT RESPONSE_duplicated_79" ~ 0,
    is.na(response) ~ 0
))
```



### 4. Making list with response times

Here the process data used is pre-processed and made compatible with the item responses.

```{r}
# Further on response time to merge. 
timss.speed <- timss.speed.proto
resp.items <- colnames(timss.speed$France[,6:length(timss.speed$France)])

# Numeric times
for(i in 1:length(timss.speed)){
  timss.speed[[i]][,6:33] <- sapply(timss.speed[[i]][,6:33], function(x){as.numeric(as.character(x))})
}

# Pivot longer response time data frame
timss.speed <- lapply(timss.speed, pivot_longer, all_of(resp.items), names_to = "item", values_to = "time")

# renaming items
for(i in 1:length(timss.speed)){
timss.speed[[i]]$item <- gsub("_S", "", timss.speed[[i]]$item)
}

# Making subitems of items and renaming item for merge
timss <- lapply(timss, mutate, "subitem" = item)
for(i in 1:length(timss)){
  timss[[i]][,6] <- timss[[i]] %$% gsub("B", "", item)
  timss[[i]][,6] <- timss[[i]] %$% gsub("A", "", item)
}
```


### 5. Joining lists of response times and scores

The item responses and process data are joined together.

```{r warning=FALSE, message=FALSE}
# Final frame, joining
big.timss <- list(NA)
for(i in 1:7){
big.timss[[i]] <- full_join(timss[[i]], timss.speed[[i]])
}

names(big.timss) <- names

# 1 big df
timss.df <- rbind(big.timss[[1]], big.timss[[2]], big.timss[[3]], big.timss[[4]], big.timss[[5]], big.timss[[6]], big.timss[[7]])
```


### 6. Adding data to dexter

The data is added to a data base to make it possible that classical psychometric properties can be investigated.

```{r}
# Making scoring rules
scoring_rules <- timss.df %>% select(subitem, response, item_score) %>% unique()

colnames(scoring_rules) <- c("item_id", "response", "item_score")

# Renaming timss.df columns for dexter
timss.df <- rename(timss.df, c("subitem" = "item_id", "IDBOOK" = "booklet_id", "IDSTUD" = "person_id"))

# Creating unique person_id (person id is in multiple countries ...)
timss.df$person_id <- paste(timss.df$IDCNTRY, timss.df$person_id, sep = "")

# Relevant columns
dexter.timss <- timss.df %>% select(item_id, response, person_id, booklet_id)
  
# Adding to dexter
timss.db <- start_new_project(scoring_rules, "timss.db")
add_response_data(timss.db, dexter.timss)

# Tia table
tia_tables(timss.db)
tia_tables(timss.db, )
```


### 7. Person data added - First DIF

```{r Adding person properties}
# Getting properties
person.props <- timss.df %>% select(ITSEX, person_id, IDCNTRY) %>% unique

# Renaming countries
person.props$IDCNTRY <- recode(person.props$IDCNTRY, "250" = "FRA", "926" = "ENG", "344" = "HKG", "410" = "KOR", "578" = "NOR", "634" = "QAT", "158" = "TWN")

timss.df$IDCNTRY <- recode(timss.df$IDCNTRY, "250" = "FRA", "926" = "ENG", "344" = "HKG", "410" = "KOR", "578" = "NOR", "634" = "QAT", "158" = "TWN")


# Adding properties
add_person_properties(timss.db, person.props)

table <- tia_tables(timss.db)

tia_tables(timss.db, IDCNTRY == "FRA")
```


### 8. Plotting pvalues per country

Item responses are plotted to investigate a first indication for possible DIF items.

```{r warning=FALSE}
# Relevant variables
test.df <- timss.df %>% select(IDCNTRY, item_id, item_score)

# Getting mean scores per country to see possible DIF
mean.per.country <- aggregate(test.df$item_score, by = list(timss.df$IDCNTRY, timss.df$item_id), FUN = mean, na.rm=T)


# Plotting
mean.per.country %>% filter(Group.1 == "FRA" | Group.1 == "TWN" | Group.1 == "QAT" | Group.1 == "ENG" | Group.1 == "KOR") %>% ggplot(aes(x = reorder(x, Group.2),y = Group.2)) +
  geom_point(aes(colour = Group.1))

mean.per.country %>%  ggplot(aes(x = x, y = Group.2)) +
  geom_point(aes(colour = Group.1))


test.df %>% 
  group_by(IDCNTRY, item_id) %>%
  dplyr::summarise(m = mean(item_score, na.rm = T)) %>% 
  ggplot(., aes(x = reorder(item_id, m), y = m, col = IDCNTRY))+
  geom_point()+
  coord_flip()
  
```


### 9. Head of data set

```{r}
head(timss.df)
```

### 10. Saving data sets

```{r}
# Making country dummies

timss.df$dum.fr <- ifelse(timss.df$IDCNTRY == "FRA", 1, 0)
timss.df$dum.eng <- ifelse(timss.df$IDCNTRY == "ENG", 1, 0)
timss.df$dum.hk <- ifelse(timss.df$IDCNTRY == "HKG", 1, 0)
timss.df$dum.kor <- ifelse(timss.df$IDCNTRY == "KOR", 1, 0)
timss.df$dum.qat <- ifelse(timss.df$IDCNTRY == "QAT", 1, 0)
timss.df$dum.nor <- ifelse(timss.df$IDCNTRY == "NOR", 1, 0)
timss.df$dum.twn <- ifelse(timss.df$IDCNTRY == "TWN", 1, 0)


# To long wide format for mimic

timss.df.long <- timss.df %>% dplyr::select(-response, -item, -time) %>% pivot_wider(names_from = item_id, values_from = item_score) 

# Time df
times.df.long <- timss.df %>% dplyr::select(-response, -item_score, - item)
time_id <- paste(unique(timss.df$item_id), "t", sep = "")
times.df.long$item_id2 <- rep(time_id, 1951)
times.df.long <- times.df.long %>% dplyr::select(-item_id)

times.df.long <- times.df.long %>% pivot_wider(names_from = item_id2, values_from = time)

# join to good frame 
mimic.df <-left_join(timss.df.long, times.df.long) 

# countries as names
mimic.df$IDCNTRY <- recode(mimic.df$IDCNTRY, "250" = "FRA", "926" = "ENG", "344" = "HKG", "410" = "KOR", "578" = "NOR", "634" = "QAT", "158" = "TWN")

# save df with raw response times
write.csv(mimic.df, "mimic_df_raw_times.csv")
```


```{r}
# Centering times
times.df <- mimic.df[,45:75]
mean.matrix <- matrix(NA, nrow = nrow(times.df), ncol = ncol(times.df))

for(i in 1:ncol(times.df)){
  mean.matrix[,i] <- apply(times.df[,-i], 1, mean)
}

times.std <- times.df - mean.matrix
mimic.df.stand.t <- mimic.df
mimic.df.stand.t[,45:75] <- times.std

# Save df with centered response times
write.csv(mimic.df.stand.t, "mimic_df_standtimes.csv")

```


```{r}
# Adding continent
timss.cont <- timss.df.long %>% filter(IDCNTRY != "QAT")
timss.cont$cont <- ifelse(timss.cont$IDCNTRY == "FRA" | timss.cont$IDCNTRY == "ENG" | timss.cont$IDCNTRY == "NOR", 1, 0)

# So 1 for Europe, 0 for asia
cont.times <- mimic.df.stand.t %>% filter(IDCNTRY != "QAT")
cont.times$cont <- ifelse(cont.times$IDCNTRY == "FRA" | cont.times$IDCNTRY == "ENG" | cont.times$IDCNTRY == "NOR", 1, 0)

# Save df for optional analyses with continents
write.csv(cont.times, "mimic_cont_explain.csv" )
write.csv(timss.cont, "mimic_cont_detect.csv")
```


### 11. Session information

```{r}
# Sessioninfo
sessionInfo()
```

